quollr: An R Package for Visualizing 2-D Models from Non-linear Dimension Reductions in High Dimensional Space

### 2025-11-30 paper-quollr.tex
Added the current version

\subsection{Scaling the data}\label{scaling-the-data}

@@

\subsection{Computing hexagon grid configuration}\label{computing-hexagon-grid-configuration}

The function \texttt{calc\_bins\_y()} determines the configuration of the hexagon grid by computing the number of bins along the y-axis (\texttt{b2}), the hexagon binwidth (\texttt{a1}), and binheight (\texttt{a2}). This function accepts (i) an object (\texttt{nldr\_scaled\_obj}) containing three elements: the first is the scaled NLDR data (\texttt{scaled\_nldr}), and the second and third are the limits of the original NLDR data (\texttt{lim1} and \texttt{lim2}); (ii) the number of bins along the x-axis (\texttt{b1}), and (iii) the buffer amount as a proportion (\texttt{q}). The buffer ensures that the grid fully covers the data space by extending one hexagon width (\(a_1\)) and height (\(a_2\)) beyond the observed data in all directions. By default, \(q = 0.1\), but it must be set to a value smaller than the minimum data value to avoid exceeding the data range.

@@

\subsection{Binning the data}\label{binning-the-data}

@@

\subsubsection{Mapping the points to their corresponding hexagonal bins}\label{mapping-the-points-to-their-corresponding-hexagonal-bins}

\subsection{Obtaining bin centroids}\label{obtaining-bin-centroids}

@@

\subsection{Indicating neighbors by line segments connecting centroids}\label{indicating-neighbors-by-line-segments-connecting-centroids}

To represent the neighborhood structure of hexagonal bins in a \(2\text{-}D\) layout, we employ Delaunay triangulation \citep{lee1980, albrecht2024} on the centroids of hexagons. This geometric approach is used to infer which bins are considered neighbors.

The \texttt{tri\_bin\_centroids()} function generates a triangulation object from the x and y coordinates of hexagon centroids using the \texttt{interp::tri.mesh()} function \citep{albrecht2024}. This triangulation forms the structural basis for identifying adjacent bins.

@@

The \texttt{gen\_edges()} function uses this triangulation object to extract line segments between neighboring bins. It constructs a unique set of bin-to-bin connections by identifying the triangle edges and filtering duplicate or reversed links. Each edge is then annotated with its start and end coordinates, and a Euclidean distance is computed using the helper function \texttt{calc\_2d\_dist()}.

>>>>

@@

\subsubsection{Hexagon binning}\label{hexagon-binning}

The function \texttt{calc\_bins\_y()} determines the configuration of the hexagon grid by computing the number of bins along the y-axis (\texttt{b2}), the hexagon binwidth (\texttt{a1}), and hexagon binheight (\texttt{a2}). This function accepts (i) an object (\texttt{nldr\_scaled\_obj}) containing three elements: the first is the scaled NLDR data (\texttt{scaled\_nldr}), and the second and third are the limits of the original NLDR data (\texttt{lim1} and \texttt{lim2}); (ii) the number of bins along the x-axis (\texttt{b1}), and (iii) the buffer amount as a proportion (\texttt{q}). The buffer ensures that the grid fully covers the data space by extending one hexagon width (\(a_1\)) and height (\(a_2\)) beyond the observed data in all directions. By default, \(q = 0.1\), but it must be set to a value smaller than the minimum data value to avoid exceeding the data range.

@@

@@

\subsubsection{Obtaining bin centroids}\label{obtaining-bin-centroids}

@@

\subsubsection{Indicating neighbors by line segments connecting centroids}\label{indicating-neighbors-by-line-segments-connecting-centroids}

To represent the neighborhood structure of hexagonal bins in a \(2\text{-}D\) layout, we employ Delaunay triangulation \citep{lee1980, albrecht2024} on the centroids of hexagons.

The \texttt{tri\_bin\_centroids()} function generates a triangulation object from the x and y coordinates of hexagon centroids using the \texttt{interp::tri.mesh()} function \citep{albrecht2024}.

@@

The \texttt{gen\_edges()} function uses this triangulation object to extract line segments between neighboring bins. It constructs a unique set of bin-to-bin connections by identifying the triangle edges and filtering duplicate or reversed links. Each edge is then annotated with its start and end coordinates, and a Euclidean distance between the coordinates.
